Continous time Markov chain models; \\ extension, simulation and visualisation

### 2014-11-18 appendices/appendixb.tex
summary
B-W figures for appendices

\chapter{Tutorial on using SPREAD software}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig01}
\par\end{centering}
\caption{Loading MCC tree file.}
\label{fig:01}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig02}
\caption{Message in Terminal.}
\label{fig:02}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig003}
\caption{Location coordinates editor.}
\label{fig:003}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig0003}
\caption{Terminal output.}
\label{fig:0003}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig03}
\caption{Color chooser.}
\label{fig:03}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig04}
\caption{Message in Terminal.}
\label{fig:04}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{H5N1_discrete_vis}
\caption{Screenshot of the Spread output.}
\label{fig:05}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig06}
\caption{Loading the log file.}
\label{fig:06}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig006}
\caption{Terminal output.}
\label{fig:006}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{H5N1_bf_vis}
\caption{Generated visualisations.}
\label{fig:07}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.2]{fig08}
\caption{Terminal tab output.}
\label{fig:08}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig09}
\caption{Loading tree file.}
\label{fig:09}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{RacRABV_cont_vis}
\caption{TScreenshot of the Spread output.}
\label{fig:RacRABV_cont_vis}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.2]{fig10}
\caption{Choosing analysis type.}
\label{fig:10}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{RacRABV_Time_Slicer_vis}
\caption{Terminal tab output.}
\label{fig:12}
\par\end{centering}
\end{figure}

>>>>

\chapter{Tutorial on using SPREAD software\label{app:spread_tuto}}
\chaptermark{Tutorial: SPREAD software}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig01}
\caption{
{ \footnotesize 
{\bf Loading MCC tree file.}
} % END: footnotesize
}
\label{fig:01}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig02}
\caption{
{ \footnotesize 
{\bf Message in Terminal.}
} % END: footnotesize
}
\label{fig:02}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig003}
\caption{
{ \footnotesize 
{\bf Location coordinates editor.}
} % END: footnotesize
}
\label{fig:003}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig0003}
\caption{
{ \footnotesize 
{\bf Terminal output.}
} % END: footnotesize
}
\label{fig:0003}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig03}
\caption{
{ \footnotesize 
{\bf Color chooser.}
} % END: footnotesize
}
\label{fig:03}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig04}
\caption{
{ \footnotesize 
{\bf Message in Terminal}
} % END: footnotesize
}
\label{fig:04}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{H5N1_discrete_vis}
\caption{
{ \footnotesize 
{\bf Screenshot of the Spread output.}
} % END: footnotesize
}
\label{fig:05}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig06}
\caption{
{ \footnotesize 
{\bf Loading the log file.}
} % END: footnotesize
}
\label{fig:06}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig006}
\caption{
{ \footnotesize 
{\bf Terminal output.}
} % END: footnotesize
}
\label{fig:006}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{H5N1_bf_vis}
\caption{
{ \footnotesize 
{\bf Generated visualisations.}
} % END: footnotesize
}
\label{fig:07}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.2]{fig08}
\caption{
{ \footnotesize 
{\bf Terminal tab output.}
} % END: footnotesize
}
\label{fig:08}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{fig09}
\caption{
{ \footnotesize 
{\bf Loading tree file.}
} % END: footnotesize
}
\label{fig:09}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{RacRABV_cont_vis-bw}
\caption{
{ \footnotesize 
{\bf Screenshot of the Spread output.}
} % END: footnotesize
}
\label{fig:RacRABV_cont_vis}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.2]{fig10}
\caption{
{ \footnotesize 
{\bf Choosing analysis type.}
} % END: footnotesize
}
\label{fig:10}
\par\end{centering}
\end{figure}

@@

\begin{figure}[H]
\begin{centering}
\includegraphics[scale=0.25]{RacRABV_Time_Slicer_vis}
\caption{
{ \footnotesize 
{\bf Terminal tab output.}
} % END: footnotesize
}
\label{fig:12}
\par\end{centering}
\end{figure}

### 2014-11-18 appendices/appendixc.tex
summary
B-W figures for appendices

\chapter{Tutorial to parallel BEAST/BEAGLE utility for sequence simulation ({\bussname})}

@@

In order to simulate character sequences, one will need a tree topology, with correspondent branch lengths, as  backbone. In {\bussname}, to import a tree, the user needs simply to go the \textbf{Trees} tab and click the `Choose file' button. After that, a window will open for you to browse though your files and go the desired tree file (see Figure~\ref{fig:loadingtopology}). {\bussname} accepts input trees in newick or nexus formats.
%%
\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{LOAD_TREE.png} 
\end{center}
\caption{How to load a tree into {\bussname}}
\label{fig:loadingtopology}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{DEMOGRAPHIC.png} 
\end{center}
\caption{Choosing a demographic (coalescent) model}
\label{fig:demographic}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{HKY.png} 
\end{center}
\caption{Setting a HKY substitution model with $\kappa$ parameter of 1.0.}
\label{fig:hky}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{GY94.png} 
\caption{Setting a codon model (GY94) with $\kappa=1.0$ and $\omega=0.1$.}
\label{fig:gy94}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{GAMMA.png} 
\caption{Site heterogeinity can be simulated by going to the \textbf{Partitions} and click on the `Site Rate Model' button.}
\label{fig:gamma}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{BASE_FREQUENCIES.png} 
\caption{To specify base frequencies, go to the \textbf{Partitions} table and click the `Frequency Model' button.}
\label{fig:bases}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{CLOCK2.png} 
\caption{ Setting up an Inverse Gaussian relaxed clock with mean$=0$, unit standard deviation and no offset}
\label{fig:clock2}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{MULTIPLE_PARTITIONS.png} 
\caption{You can set up any number of partitions, each with its own tree topology, model of evolution, molecular clock and base frequency. Just click the `+' button at the bottom left corner of the GUI window at the \textbf{Partitions} tab.}
\label{fig:partitions}
\end{figure}
%%

@@

% FB: how to color cells in multirow table?
% TODO: LM: Update the table to show AA options
\begin{table}[H]
\begin{center}
\footnotesize{
\begin{tabular}{lcc}
\hline 
\textbf{Parameter} & \textbf{Command} & \textbf{Options}\tabularnewline
\hline 
\cellcolor{snow3}Tree topology & \cellcolor{snow3}-treeFile & \cellcolor{snow3}-\tabularnewline
\hline 
Taxa set & -taxaSet & -\tabularnewline
\hline 
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-constantPopulationParameterValues\tabularnewline
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-exponentialGrowthRateParameterValues\tabularnewline
\cellcolor{snow3}\multirow{-3}{*}{Demographic (coalescent) model} & \cellcolor{snow3}\multirow{-3}{*}{-demographicModel} & \cellcolor{snow3}-exponentialDoublingTimeParameterValues\tabularnewline
\hline 
 &  & -HKYsubstitutionParameterValues\tabularnewline
 &  & -GTRsubstitutionParameterValues\tabularnewline
 &  & -TN93substitutionParameterValues\tabularnewline
\multirow{-4}{*}{Markov model of evolution} & \multirow{-4}{*}{-branchSubstitutionModel} & -GY94substitutionParameterValues\tabularnewline
\hline 
\cellcolor{snow3}Site rate heterogeinity & \cellcolor{snow3}-siteRateModel & \cellcolor{snow3}-gammaSiteRateModelParameterValues\tabularnewline
\hline 
 &  & -strictClockParameterValues\tabularnewline
 &  & -lognormalRelaxedClockParameterValues\tabularnewline
 &  & -exponentialRelaxedClockParameterValues\tabularnewline
\multirow{-4}{*}{Molecular clock model} & \multirow{-4}{*}{-clockRateModel} & -inverseGaussianRelaxedClockParameterValues\tabularnewline
\hline 
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-nucleotideFrequencyParameterValues\tabularnewline
\cellcolor{snow3}\multirow{-2}{*}{Base frequencies} & \cellcolor{snow3}\multirow{-2}{*}{-baseFrequencies} & \cellcolor{snow3}-codonFrequencyParameterValues\tabularnewline
\hline 
Root sequence & -rootSequence & -\tabularnewline
\end{tabular}
} % END: footnotesize
\caption{{ \footnotesize
List of command line options for {\bussname}.
}} 
\label{tab:commands}
\end{center}
\end{table}

@@

Using the CLI, seed is parsed after parsing all the arguments for partitions.
The user can specify a seed and choose to use the parallel implementation by adding 
\begin{code} % TODO: rather not add it to the table, see if you agree
[<output-file-name>] [<seed>] [<true|false>] [<true|false>]
\end{code}
after the last ``:'' sign dividing the desired partitions.
The first true/false statement allows the ancestral sequences in the inner nodes to  be saved -- also accomplished by checking the corresponding box in the {\bf Simulation} tab.
The last argument corresponds to the use of the parallel implementation, which is false by default. 
%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{SIMUTAB.png} 
\caption{To set the seed for your simulation and choose the parallel BEAGLE implementation, check the corresponding tick boxes at the {\bf Simulation} tab.}
\label{fig:simutab}
\end{figure}
%%
\subsection{Generating data evolving under GY94 codon model}
To generate sequence data evolving according to codon substitution model we start by importing a backbone tree topology. 
Load the \href{http://rega.kuleuven.be/cev/ecv/software/buss_files/simtree.figtree}{\emph{SimTree.figtree}} tree file shipped with this tutorial as described in \emph{\loading} subsection.

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{CODON_FREQUENCIES.png} 
\end{center}
\caption{
Setting codon frequencies. Just select codon frequencies in the drop-down menu and you're done. The default is $1/61$ for each codon.
} % END: caption
\label{fig:CodonFrequencies}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{CODON_PARTITIONING.png} 
\end{center}
\caption{
Codon partitioned simulation.
} % END: caption
\label{fig:CodonPartitoning}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{TAXA_EDITOR.png} 
\end{center}
\caption{
Loading a taxa set from tab-separated file. Once you load the taxa set onto {\bussname}, you are ready to simulate under any of the coalescent dynamics available. 
} % END: caption
\label{fig:TaxaEditor}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{DEMOGRAPHIC_MODEL.png} 
\end{center}
\caption{
Loading a taxa set from tab-separated file.
} % END: caption
\label{fig:DemoModel}
\end{figure}

@@

\begin{figure}[H]
\begin{center}
\includegraphics[scale=\figurescale]{TERMINAL_OUTPUT.png} 
\end{center}
\caption{
Terminal output
} % END: caption
\label{fig:Terminal}
\end{figure}

@@

We need to define the discrete geographical locations that {\bussname} will sample from. After the \emph{taxa} block paste the following code:

@@

Scroll to the \emph{frequencyModel} block. Because we have specified $K=3$ discrete locations we need to specify base frequencies for them. Delete the existing \emph{frequencyModel} block and paste there following code instead:

@@

Now scroll further down to the substitution model block (defined between \emph{HKYModel} tags if you generated the XML using default settings). We need to define a geographical substitution model, driven by a rate matrix. 
We specify $K\times(K-1)=6$ entries such that the underlying instantaneous rate matrix is asymmetric (non-reversible). 
Replace the existing block with:

@@

Now reference the created substitution model inside \emph{siteModel} tags like this:

@@

Finally all the new elements need to be referenced in the \emph{beagleSequenceSimulator} block. We specify one replicate, as every taxa can ultimately have only one geographical trait:
\begin{lstlisting}
<beagleSequenceSimulator id="simulator">
  <partition from="1" to="1"> 
    <treeModel idref="treeModel1"/>
    <complexSubstitutionModel idref="originModel.simulation"/>	    
    <siteModel idref="geoSiteModel.simulation"/>
    <strictClockBranchRates idref="branchRates1"/>
    <frequencyModel idref="freqModel"/>
  </partition>
</beagleSequenceSimulator>
\end{lstlisting}

>>>>

\chapter{Tutorial to parallel BEAST/BEAGLE utility for sequence simulation ({\bussname})\label{app:pibuss_tuto}}
\chaptermark{Tutorial: {\bussname}}

@@

In order to simulate character sequences, one will need a tree topology, with correspondent branch lengths, as  backbone. In {\bussname}, to import a tree, the user needs simply to go the \textbf{Trees} tab and click the `Choose file' button. After that, a window will open for you to browse though your files and go the desired tree file (see Figure~\ref{fig:loadingtopology}). {\bussname} accepts input trees in newick or nexus formats.

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{LOAD_TREE.png} 
\caption{
{ \footnotesize 
{\bf How to load a tree into {\bussname}.}
} % END: footnotesize
}
\label{fig:loadingtopology}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{DEMOGRAPHIC.png} 
\caption{
{ \footnotesize 
{\bf Choosing a demographic (coalescent) model.}
} % END: footnotesize
}
\label{fig:demographic}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{HKY.png} 
\caption{
{ \footnotesize 
{\bf Setting a HKY substitution model with $\kappa$ parameter of $1.0$.}
} % END: footnotesize
}
\label{fig:hky}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{GY94.png} 
\caption{
{ \footnotesize 
{\bf Setting a codon model (GY94) with $\kappa=1.0$ and $\omega=0.1$.}
} % END: footnotesize
}
\label{fig:gy94}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{GAMMA.png} 
\caption{
{ \footnotesize 
{\bf Site rate heterogeinity.}
Site heterogeinity can be simulated by going to the \textbf{Partitions} and click on the `Site Rate Model' button.
} % END: footnotesize
}
\label{fig:gamma}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{BASE_FREQUENCIES.png} 
\caption{
{ \footnotesize 
{\bf Base frequencies.}
To specify base frequencies, go to the \textbf{Partitions} table and click the `Frequency Model' button.
} % END: footnotesize
}
\label{fig:bases}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{CLOCK2.png} 
\caption{
{ \footnotesize 
{\bf Clock model.}
Setting up an Inverse Gaussian relaxed clock with mean$=0$, standard deviation of $1.0$ and no offset.
} % END: footnotesize
}
\label{fig:clock2}
\end{figure}
%%

@@

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{MULTIPLE_PARTITIONS.png} 
\caption{
{ \footnotesize 
{\bf Setting multiple partitions.}
You can set up any number of partitions, each with its own tree topology, model of evolution, molecular clock and base frequency. 
Just click the `+' button at the bottom left corner of the GUI window at the \textbf{Partitions} tab.
} % END: footnotesize
}
\label{fig:partitions}
\end{figure}
%%

@@

% FB: how to color cells in multirow table?
% TODO: LM: Update the table to show AA options
\begin{table}[H]
\begin{center}
\footnotesize{
\begin{tabular}{lcc}
\hline 
\textbf{Parameter} & \textbf{Command} & \textbf{Options}\tabularnewline
\hline 
\cellcolor{snow3}Tree topology & \cellcolor{snow3}-treeFile & \cellcolor{snow3}-\tabularnewline
\hline 
Taxa set & -taxaSet & -\tabularnewline
\hline 
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-constantPopulationParameterValues\tabularnewline
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-exponentialGrowthRateParameterValues\tabularnewline
\cellcolor{snow3}\multirow{-3}{*}{Demographic (coalescent) model} & \cellcolor{snow3}\multirow{-3}{*}{-demographicModel} & \cellcolor{snow3}-exponentialDoublingTimeParameterValues\tabularnewline
\hline 
 &  & -HKYsubstitutionParameterValues\tabularnewline
 &  & -GTRsubstitutionParameterValues\tabularnewline
 &  & -TN93substitutionParameterValues\tabularnewline
\multirow{-4}{*}{Markov model of evolution} & \multirow{-4}{*}{-branchSubstitutionModel} & -GY94substitutionParameterValues\tabularnewline
\hline 
\cellcolor{snow3}Site rate heterogeinity & \cellcolor{snow3}-siteRateModel & \cellcolor{snow3}-gammaSiteRateModelParameterValues\tabularnewline
\hline 
 &  & -strictClockParameterValues\tabularnewline
 &  & -lognormalRelaxedClockParameterValues\tabularnewline
 &  & -exponentialRelaxedClockParameterValues\tabularnewline
\multirow{-4}{*}{Molecular clock model} & \multirow{-4}{*}{-clockRateModel} & -inverseGaussianRelaxedClockParameterValues\tabularnewline
\hline 
\cellcolor{snow3} & \cellcolor{snow3} & \cellcolor{snow3}-nucleotideFrequencyParameterValues\tabularnewline
\cellcolor{snow3}\multirow{-2}{*}{Base frequencies} & \cellcolor{snow3}\multirow{-2}{*}{-baseFrequencies} & \cellcolor{snow3}-codonFrequencyParameterValues\tabularnewline
\hline 
Root sequence & -rootSequence & -\tabularnewline
\end{tabular}
} % END: footnotesize
\caption{
{ \footnotesize 
{\bf List of command line options for {\bussname}.}
} % END: footnotesize
}
\label{tab:commands}
\end{center}
\end{table}

@@

Using the CLI, seed is parsed after parsing all the arguments for partitions.
The user can specify a seed and choose to use the parallel implementation by adding 
\begin{code} % TODO: rather not add it to the table, see if you agree
[<output-file-name>] [<seed>] [<true|false>] [<true|false>]
\end{code}
after the last ``:'' sign dividing the desired partitions.
The first true/false statement allows the ancestral sequences in the inner nodes to  be saved -- also accomplished by checking the corresponding box in the {\bf Simulation} tab.
The last argument corresponds to the use of the parallel implementation, which is false by default.

%%
\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{SIMUTAB.png} 
\caption{
{ \footnotesize 
{\bf Setting RNG seed.}
To set the seed for your simulation and choose the parallel BEAGLE implementation, check the corresponding tick boxes at the {\bf Simulation} tab.
} % END: footnotesize
}
\label{fig:simutab}
\end{figure}
%%

\subsection{Generating data evolving under GY94 codon model}

To generate sequence data evolving according to codon substitution model we start by importing a backbone tree topology. 
Load the \href{http://rega.kuleuven.be/cev/ecv/software/buss_files/simtree.figtree}{\emph{SimTree.figtree}} tree file shipped with this tutorial as described in \emph{\loading} subsection.

@@

\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{CODON_FREQUENCIES.png} 
\caption{
{ \footnotesize 
{\bf Setting codon frequencies.}
Just select codon frequencies in the drop-down menu and you're done. The default is $1/61$ for each codon.
} % END: footnotesize
}
\label{fig:CodonFrequencies}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{CODON_PARTITIONING.png} 
\caption{
{ \footnotesize 
{\bf Codon partitioned simulation.}
} % END: footnotesize
}
\label{fig:CodonPartitoning}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{TAXA_EDITOR.png} 
\caption{
{ \footnotesize 
{\bf Loading a taxa set from tab-separated file.}
Once you load the taxa set onto {\bussname}, you are ready to simulate under any of the coalescent dynamics available.
} % END: footnotesize
}
\label{fig:TaxaEditor}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{DEMOGRAPHIC_MODEL.png} 
\caption{
{ \footnotesize 
{\bf Loading a taxa set from tab-separated file.}
} % END: footnotesize
}
\label{fig:DemoModel}
\end{figure}

@@

\begin{figure}[H]
\centering
\includegraphics[scale=\figurescale]{TERMINAL_OUTPUT.png} 
\caption{
{ \footnotesize 
{\bf Terminal output.}
} % END: footnotesize
}
\label{fig:Terminal}
\end{figure}

@@

We need to define the discrete geographical locations that {\bussname} will sample from. After the  {\color{darkblue}taxa} block paste the following code:

@@

Scroll to the {\color{darkblue}frequencyModel} block. Because we have specified $K=3$ discrete locations we need to specify base frequencies for them. Delete the existing {\color{darkblue}frequencyModel} block and paste there following code instead:

@@

Now scroll further down to the substitution model block (defined between {\color{darkblue}HKYModel} tags if you generated the XML using default settings). 
We need to define a geographical substitution model, driven by a rate matrix. 
We specify $K\times(K-1)=6$ entries such that the underlying instantaneous rate matrix is asymmetric (non-reversible). 
Replace the existing block with:

@@

Now reference the created substitution model inside {\color{darkblue}siteModel} tags like this:

@@

Finally all the new elements need to be referenced in the {\color{darkblue}beagleSequenceSimulator} block. We specify one replicate, as every taxa can ultimately have only one geographical trait:

\begin{lstlisting}
<beagleSequenceSimulator id="simulator">
  <partition from="1" to="1"> 
    <treeModel idref="treeModel1"/>
    <complexSubstitutionModel idref="originModel.simulation"/>	    
    <siteModel idref="geoSiteModel.simulation"/>
    <strictClockBranchRates idref="branchRates1"/>
    <frequencyModel idref="freqModel"/>
  </partition>
</beagleSequenceSimulator>
\end{lstlisting}

### 2014-11-18 chapters/chapter06.tex
B-W figures for appendices

\begin{figure}
\centering
$$\bullet\; C\;\star C\;\star\; A\;\star\; C\;\star\; T\;\star$$
\caption{
{ \footnotesize 
{\bf DNA sequence CCACT of length $L=5$, with $L+1$ links that separate the characters.} 
By convention to the right of each base there is a \textit{normal link} ($\star$); the leftmost base has an \textit{immortal link} to itself($\bullet$), ensuring that an insertion can occur at position $l=0$.
} % END: footnotesize
}
\label{fig:indels}
\end{figure}

@@

Traditionally a vast majority of programs were written as sequential code, following von Neumanns design, which proved to be working well for the scientific community for many years.
This relates to an observation, which later came to be known as Moore's law, predicting the trend in chip performance as doubling roughly every $18$ months.
%As captured by an observation that later came to be known as Moore's law, chip performance was predicted to double roughly every $18$ months.
Figure~\ref{fig:moore} shows how, during the time of $15$ years, the hardware performance and improvements in compiler technology provided a speedup by a factor of $80$.
%PL: whats the best and average in the plot; was it taken from somewhere -- in other words does it needs acknowledgements?
For practical purposes this meant that just by updating the hardware architecture and compilers increasingly bigger problems could be tackled, without ever changing the code-base of the numerical programs. 
The same software would simply run faster with the introduction of new generations of processors and new versions of compilers.
%This has created a positive cycle for the computer industry.
However, during the last two years in the plot this gain is no longer visible\footnote{The free ride is over!}.
Due to hardware limitations, and problems with heat dissipation, we are now at the limit of how many transistors can be put into integrated circuits, which determines CPU performance.
As a consequence, serial computations on new CPUs no longer benefit from the same increases in speed and performance.

\begin{figure}[H]
\centering
\includegraphics[scale=0.2]{moore}
\caption{
{ \footnotesize 
{\bf Performance gains of a numerical code over the course of 15 years.}
%PL: whats the best and average in the plot; was it taken from somewhere -- in other words does it needs acknowledgements?
} % END: footnotesize
}
\label{fig:moore}
\end{figure}

>>>>

\begin{figure}
\centering
$$\bullet\; C\;\star C\;\star\; A\;\star\; C\;\star\; T\;\star$$
\caption{
{ \footnotesize 
{\bf DNA sequence CCACT of length $L=5$, with $L+1$ links that separate the characters.} 
By convention to the right of each base there is a \textit{normal link} ($\star$); the leftmost base has an \textit{immortal link} to itself ($\bullet$), ensuring that an insertion can occur at position $l=0$.
} % END: footnotesize
}
\label{fig:indels}
\end{figure}

@@

Traditionally a vast majority of programs were written as sequential code, following von Neumanns design, which proved to be working well for the scientific community for many years.
This relates to an observation, which later came to be known as Moore's law, predicting the trend in chip performance as doubling roughly every $18$ months.
%As captured by an observation that later came to be known as Moore's law, chip performance was predicted to double roughly every $18$ months.
Figure~\ref{fig:moore} shows how, during the time of $15$ years, the hardware performance and improvements in compiler technology provided a speedup by a factor of $80$.
For practical purposes this meant that just by updating the hardware architecture and compilers increasingly bigger problems could be tackled, without ever changing the code-base of the numerical programs. 
The same software would simply run faster with the introduction of new generations of processors and new versions of compilers.
%This has created a positive cycle for the computer industry.
However, during the last two years in the plot this gain is no longer visible\footnote{The free lunch is over!}.
Due to hardware limitations, and problems with heat dissipation, we are now at the limit of how many transistors can be put into integrated circuits, which determines CPU performance.
As a consequence, serial computations on new CPUs no longer benefit from the same increases in speed and performance.

\begin{figure}[H]
\centering
\includegraphics[scale=0.2]{moore}
\caption{
{ \footnotesize 
{\bf Performance gains of a numerical code benchmarked on a single CPU core over the course of 15 years \citep{Turek1999}.}
Last two years have seen only a marginal speed-up, and correspond to the shift to multicore architectures.
} % END: footnotesize
}
\label{fig:moore}
\end{figure}
